Improved Bounds for Online Balanced Repartitioning

### 2020-03-29 paper/paper.tex
todos

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
    \rowcolor{gray!50}
    \textbf{Clusters/capacity} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 3\hspace{0.3cm}(\S 9.3) \\ 
      \rowcolor{gray!25}
      \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
      \textbf{$\ell=2$}&  $2k-1$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
      \rowcolor{gray!25}
      general $k$ and $\ell$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
    \end{tabular}
                \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound.
                  }
  \label{tab:overview}
  \vspace{-7mm}
\end{table*}

@@

\noindent \textbf{Preliminaries.}
In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes.
We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
For $k=2$ a constant competitive algorithm exists.
At the end of this section, we also
discuss an upper bound.

\maciek{TODO: better introduction of the concept of adversary's components in the lower bound. Perhaps other name than adversary components to avoid confusion.}

@@

@@

\begin{proof}
	We construct an instance of the problem with $\ell$ clusters 
	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
	Fix any online algorithm \ALG{}.
  We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
	If \ALG{} at any point splits a component
	(i.e., spreads its nodes over two or more clusters),
	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
	If \ALG{} never collocates them, then it is not competitive.

@@

	Given this sequence of requests,
	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
	Hence,
	OPT pays for $2$ node swaps and
	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
  therefore $|X| \geq k \cdot \ell - k - \ell$.
  Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}

@@

\subsection{Upper bound.}
Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to a perfect partition
at the beginning and stays there perpetually.
\maciek{Already mention that the LB is for that scenario, too?}
%We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs which grow as components.
%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.

We assume \OPT begins with the initial configuration
$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
$P_F = F_1, \dots, F_{\ell}$.
 The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
 \maciek{It is OK to use single-argument $\dist$ (two-arg is a must in a definition, and it is already there).}
 \maciek{We should denote distance $\Delta(P)$, and the current $\Delta$ should be R (like radius).}
    That is,
    $\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
In other words,
at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
$\OPT \geq \Delta:= \dist(P_F, P_I) $.
 With each repartitioning,
  \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
As a result,
\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
This invariant ensures that \PPL does not pay too much while recovering $P_F$.
%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.
The scheme of the algorithm can be found in the Appendix \ref{apx:PPL}.  
\begin{property} \label{prop:dist<OPT}
    Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
    Then, $\dist(P,P_I) \leq \Delta$.
\end{property}

@@

\begin{theorem}	\label{thm:upperbound}
    \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
    \maciek{Consider not hiding the $2(k-1)\ell$ in the theorem. This is a better bound.}
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
     it does not move to a configuration
     $P', \dist(P', P_I) > \Delta$,
     before it enumerates all configurations at distance $\Delta$.
     Therefore,
     \PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
     (i.e., the number of internal edges in $P_F$).
     \maciek{Clarify why the number of internal edges bound the number of repartitions.}
    By Lemma \ref{lemma:rebalancecost},
    each repartition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
%    	So it is 4-competitive (?)}
 \end{proof}

@@

@@

  As \TAlg is component respecting, it never increases any counter above $\alpha$.
  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
  Let $\sigma$ be the input sequence that arrived during the phase.
  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.

@@

\pagebreak
\appendix
\section{PPL Algorithm} \label{apx:PPL}
\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
		\STATE{$P_0 := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
		\IF{$C_1 \neq C_2$}
		\STATE {Unite the two components into a single component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
			\COMMENT{i.e.~if not in the same cluster}    
		}       
		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
		\COMMENT{move to $P$ closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}

\section{Omitted proofs}
\label{apx:omitted}

\begin{proof}[Proof of Lemma~\ref{lem:1req}]
  Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.

@@

\maciek{TODO: splitting nodes explain in the preliminaries}

@@

Now, we claim that the cost of node migration of \OFF is $2 \cdot \alpha$ (two caching nodes are swapped) if $\OPT$ performs a fetch, and 0 if
$\OPT$ does not change its cache contents.
Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
\maciek{Finished here}
To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.

As $B$ is $\gamma$-competitive for the BRP problem, there exists a constant $\beta$, such that for any sequence $\sigma^P$ and the corresponding sequence $\sigma^B$, it holds that $B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) +
\beta$.
Combining this inequality with the inequalities between $P$ and $B$ and between \OFF and \OPT yields
\[
	2 \cdot \alpha \cdot P(\sigma^P) \leq 
	B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) + \beta
	 =\gamma \cdot 2 \cdot \alpha \cdot \OPT(\sigma^P) + \beta,
\]
and therefore $P$ is $\gamma$-competitive.
  \medskip

For any sequence of paging requests $I_{PAGE}$, we produce a corresponding BRP pairwise request sequence $I_{BRP}$.
To this end, fix an algorithm ALG for the paging problem.
When a request to element $e$ comes, we identify the set $S \in \{ A,B,C \}$ that contains the corresponding node $v_e$.
Then, the adversary forms a component $S$, and dismisses it right after the algorithm collocates the whole $S$ on one cluster.
Recall that forming a component means that the adversary continuously issues requests to any two nodes of a component that are split.

Now we lower-bound the cost of the algorithm for any sequence of pairwise requests that were produced by the described transformation of paging requests.
For any such request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to reach the state with $S$ collocated.

We construct the following offline algorithm OPT.
OPT only resides in three types of configurations with respect to sets $A,B,C$ (cf. Figure TODO).
Without loss of generality, we describe the contents of clusters $X$ and $Y$:
\begin{enumerate}
  \item In $X$ we have $A$ and one node from $C$; in $Y$ we have $B$ and the other node from $C$.
  \item In $X$ we have $A$ and one node from $B$; in $Y$ we have $C$ and $k-2$ nodes from set $B$.
  \item In $X$ we have $B$ and one node from $A$; in $Y$ we have $C$ and $k-2$ nodes from set $A$.
\end{enumerate}

Now we describe the possible actions performed by OPT.
Within each type of configuration with $S \in \{ A, B, C\}$ split, OPT may exchange the split node with any node from $S$.
To transition between any of three configuration types, it suffices to perform a single exchange.
Hence, for every action OPT pays $2\alpha$.

Finally, we describe the actions of OPT on the input sequence $I_{BRP}$.
OPT observes the actions of the algorithm LFD (Least Forward Distance for the caching problem) on $I_{PAGE}$, and performs corresponding actions within its feasible configurations.
Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split.
The cost of OPT is then $2\alpha \cdot LFD$.

It is well-known that LFD is an optimal algorithm for paging, and any deterministic algorithm for paging with cache size $2k-1$ is at least $2k-1$-competitive.

TODO: rewrite the conclusion in the similar way to sidma paper
\end{proof}

>>>>

\maciek{TODO: fill the table: results and references to sections.}

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
    \rowcolor{gray!50}
    \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 3\hspace{0.3cm}(\S 9.3) \\ 
      \rowcolor{gray!25}
      \textbf{$k=3$}&  $1357\cdot \ell - 245 = \Omega(\ell)$\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
      \textbf{$\ell=2$}&  $2k-1$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
      \rowcolor{gray!25}
      $\ell \geq 3$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
      Perfect partition & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k \cdot \ell)$\hspace{0.3cm} (\S 9) \\
    \end{tabular}
                \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound. If either $k$ or $\ell$ is missing from a variant description, it is assumed to be arbitrary.
                  }
  \label{tab:overview}
  \vspace{-7mm}
\end{table*}

@@

\noindent \textbf{Preliminaries.}

\maciek{Here, we need a good and broad description of what a component is.}

In this paper, we use the technique of maintaining \emph{(connected) components} of nodes, similar to previous approaches~\cite{repartition-disc}. A~\emph{component} is a subset of frequently communicating nodes \maciek{This is not a definition, just some intuitions. More formal needed.}.
We use this concept in both algorithms in this paper, but also in the lower bound, to incur high cost for any online algorithm that splits them.
An~algorithm is \emph{component respecting}
if it always keeps nodes that  belong to the same component collocated.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
%such that frequently communicating nodes (from different clusters)
%are relocated to the same cluster before they inflict too much communication cost.
%A \emph{reconfiguration} (i.e.,\emph{ re-partitioning})
%is performed by migrating nodes between clusters,
%at the total cost of individual node migrations.
%The objective is to jointly minimize communication and migration costs while serving the sequence.

@@

\maciek{The text before the proof should have two parts: a description of adversary components and then general intro to the proof.}

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
It is sufficient to have clusters of capacity $k\geq 3$ for this lower bound.
For $k=2$, a constant competitive algorithm exists.
At the end of this section, we also
discuss an upper bound.

\maciek{TODO: rename everywhere: "adversary's components", i.e., all components in lower bounds should be called "ground sets"}

@@

\maciek{Here: we abstract from the value of $\alpha$, but works for any alpha.}

@@

\begin{proof}
	We construct an instance of the problem with $\ell$ clusters 
	$\set{ S_1, S_2,\dots , S_{\ell}}, |S_i|  = k$.
	Let $I(C)$ denote the cluster where nodes of component $C$ are located initially.
	%	\maciek{So far components were not introduced; the best place to introduce these would be around the definition of \PPOBRP}.
  Fix any online algorithm \ALG{}.

  \maciek{This paragraph should be before the proof, merged with section about adversary components.}
  We join some nodes into components, and produce the input sequence by observing actions of \ALG{} with respect to nodes of these components.
	If \ALG{} at any point splits a component
	(i.e., spreads its nodes over two or more clusters),
	then we continue to issue requests between non-collocated nodes of the component until \ALG{} collocates the nodes of the component.
	If \ALG{} never collocates them, then it is not competitive.

@@

	Given this sequence of requests,
	the optimal strategy is to migrate $\set{x_0,y_0}$ to the cluster $I(x^*)$ by
	swapping $\set{x_0,y_0}$ with $\set{x^*,y^*}$.
	Hence,
	OPT pays for $2$ node swaps and
	\ALG{} incurs at least one swap for each node in the sequence $X := x_0, x_1,\dots$.
	We exclude at most $(k-1) + ( \ell+1)$ nodes out of $k \cdot \ell$ nodes,
  therefore $|X| \geq k \cdot \ell - k - \ell$ \maciek{elaborate on exclusion?}.
  Finally, $\ALG/\OPT \geq (k \cdot \ell - k - \ell) / 2\in \Omega(k\cdot\ell)$.
\end{proof}

@@

\subsection{Upper bound.}
Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to a perfect partition
at the beginning and stays there perpetually.
\maciek{Already mention that the LB is for that scenario}
%We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs which grow as components \maciek{rewrite sentence, expand}.
%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.

We assume \OPT begins with the initial configuration
$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
$P_F = F_1, \dots, F_{\ell}$.
 The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
 \maciek{It is OK to use single-argument $\dist$ (two-arg is a must in a definition, and it is already there).}
 \maciek{We should denote distance $\Delta(P)$, and the current $\Delta$ should be R (like radius).}
    That is,
    $\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
In other words,
at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
$\OPT \geq \Delta:= \dist(P_F, P_I) $.
 With each repartitioning,
  \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
As a result,
\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
This invariant ensures that \PPL does not pay too much while recovering $P_F$.
%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

\section{PPL Algorithm} \label{alg:PPL}
\begin{algorithm}
	\renewcommand{\algorithmicrequire}{\textbf{Input:}}
	\renewcommand{\algorithmicensure}{\textbf{Output:}}
	\begin{algorithmic}
		%        \Require 
		%        $k, \ell$,
		%        initial configuration $P_I$,
		%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
		%        \Ensure A final configuration $P_F$ 
		\STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
		\STATE{$P_0 := P_I$}
		\label{line:initcomponents}
		\FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
		\STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
		\IF{$C_1 \neq C_2$}
		\STATE {Unite the two components into a single component $C'$ and
			$\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
		\IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
			\COMMENT{i.e.~if not in the same cluster}    
		}       
		\STATE {$P_{t} = \mathit{repartition}(P_{t-1}, P_I, \mathcal{C})$} 
		\COMMENT{move to $P$ closest to $P_I$}
		\label{line:rebalance} 
		\ENDIF
		\ENDIF
		\ENDFOR
	\end{algorithmic}
	\caption{Perfect Partition Learner (\PPL)}
	\label{alg:ppl}
\end{algorithm}

\begin{property} \label{prop:dist<OPT}
    Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
    Then, $\dist(P,P_I) \leq \Delta$.
\end{property}

@@

\begin{theorem}	\label{thm:upperbound}
    \PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
    \maciek{Do not hide that we have $2(k-1)\ell$ in the theorem. This is a better bound.}
\end{theorem}
\begin{proof}
      On each inter-cluster request,
     the algorithm enumerates all $\ell$-way partitions of components
     that are in the same (closest) distance of $P_I$.
     That is, 
     once it reaches a configuration $P$ at distance $\Delta = \dist(P, P_I)$,
     it does not move to a configuration
     $P', \dist(P', P_I) > \Delta$,
     before it enumerates all configurations at distance $\Delta$.
     Therefore,
     \PPL eventually reaches $\Delta=\OPT$ and the configuration $P_F$.
%    including the request that completes revealing of all components that are collocated in $P_F$.
    There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
     (i.e., the number of internal edges in $P_F$).
     \maciek{Clarify why the number of internal edges bound the number of repartitions.}
    By Lemma \ref{lemma:rebalancecost},
    each repartition costs at most $2\cdot\OPT$.
    The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
%    	So it is 4-competitive (?)}
 \end{proof}

@@

\begin{proof}
    Observe that when the repartition is triggered by \TAlg, the resulting partitioning is component respecting.
  Otherwise, if it does not exist, \TAlg simply ends the phase and performs no repartition.

@@

  As \TAlg is component respecting, it never increases any counter above $\alpha$.
  We say that the pair $(u, v)$ is \emph{saturated} if the counter has value $\alpha$, and \emph{unsaturated} otherwise \maciek{Rename: saturated/unsaturated edge instead of a pair}.
  Let $\sigma$ be the input sequence that arrived during the phase.
  Let $\sigma_{cost}$ be the requests that at the moment of arrival were external requests for \TAlg (these are the only requests that incurred a cost for \TAlg).
  In our analysis, we partition $\sigma_{cost}$ into subsequences $\sigma_I$ and $\sigma_E$.
  The sequence $\sigma_I$ (inter-component requests) are the requests from $\sigma_{cost}$ issued to pairs that belong to the same component of \TAlg at the end of the phase.
  The sequence $\sigma_E$ (extra-component requests) denotes the requests from $\sigma_{cost}$ that do not appear in $\sigma_I$.

@@

\pagebreak
\appendix

@@

\maciek{Dont include in WAOA}

@@

Now, we claim that the cost of node migration of \OFF is $2 \cdot \alpha$ (two caching nodes are swapped) if $\OPT$ performs a fetch, and 0 if
$\OPT$ does not change its cache contents.
Within each type of configuration with $S \in \{ B, C\}$ split, $\OPT$ may achieve a configuration with a different node from $S$ outside of the cache by exchanging the split node with any node from $S$.
To transition between any of three configuration types ((1) with $A$ (2) with $B$ split and (3) with C split), it suffices to perform a single exchange.
Therefore, $\OFF(\comm_t(\sigma^P_t)) = 2 \cdot \alpha \cdot \OPT(\sigma^P_t)$, which summed over the entire sequence $\sigma^P$ yields $\OFF(\sigma^B) = 2 \cdot \alpha \cdot \OPT(\sigma^P)$.

As $B$ is $\gamma$-competitive for the BRP problem, there exists a constant $\beta$, such that for any sequence $\sigma^P$ and the corresponding sequence $\sigma^B$, it holds that $B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) +
\beta$.
Combining this inequality with the inequalities between $P$ and $B$ and between \OFF and \OPT yields
\[
	2 \cdot \alpha \cdot P(\sigma^P) \leq 
	B(\sigma^B) \leq \gamma \cdot \OFF(\sigma^B) + \beta
	 =\gamma \cdot 2 \cdot \alpha \cdot \OPT(\sigma^P) + \beta,
\]
and therefore $P$ is $\gamma$-competitive.
\end{proof}
